---
title: "Introduction to BayesTestStreak"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BayesTestStreak}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

This package provides some tools for learning about the streakiness patterns in a binary (0 and 1) sequence.

We illustrate this by looking at streaky patterns of the hitter Mike Trout for the 2016 season.

First load the package.
```{r}
library(BayesTestStreak)
```

The package contains Retrosheet play-by-play data for the 2016 season, so we can explore streaky patterns for any player that played in this season.

The `find_id` function will find the Retrosheet player id for a given player.

```{r}
trout_id <- find_id("Mike Trout")
```

The function `streak_data` will find the 0/1 streak data.  The inputs are the Retrosheet data (the data frame pbp2016 is the play-by-play data that is preloaded), the player id, and the definition of success using the EVENT_CD variable.  Here values of EVENT_CD = 20 through 23 correspond to a hit.

```{r}
y_PA <- streak_data(pbp2016, trout, 20:23)
```

Maybe you want to consider hits/no hits only among the official at-bats (instead of plate appearances).  Then you use the AB = TRUE option.

```{r}
y_AB <- streak_data(pbp2016, trout, 20:23, AB=TRUE)
```

What can you do with this data?

One can construct a moving average graph by the function `mavg_plot` -- the input is the vector of 0-1 data and the value of the width of the window where you are computing the average.

```{r}
mavg_plot(y_AB, width=40)
```

The function `find.spacings` computes the lengths of the gaps between consecutive successes.

```{r}
(s <- find.spacings(y_AB)$y)
```

The `geometric.plot` function constructs a geometric probability plot -- if the points line along the line, a geometric constant probability model is suitable.

```{r}
geometric.plot(s)
```

The `permutation.test` function tests (by simulation) the hypothesis that all possible ordering of the 0's and 1's are equally likely.

```{r}
permutation.test(y_AB)
```

The `bayes.factor.function` function implements a Bayes factor for testing for streakiness.  The graph plots the log Bayes factor against the parameter log K. (See my paper for details.)

```{r}
OUT <- bayes.factor.function(y_AB)
ggplot(data.frame(logK=OUT$log.K, logBF=OUT$log.BF), 
       aes(logK, logBF)) + geom_line()
```


